A Data-Driven Framework for Neural Field Modeling

### 2011-01-03 ltx/first_order_framework/LaTeX_Manuscript.tex
more edits to modeling assumptions

\subsection{Modeling Assumptions}
\dean{(comment 1 and 9 from reviewer 2:) It is an implicit assumption that the neural field model (in equation~\ref{DiscreteTimeModel}) provides an apt description of cortical dynamics. Although models of this form are capable of describing a variety of cortical dynamics, there is without doubt, a mismatch between the cortex and the model. Nevertheless, there is a major benefit in using a mean field model over a more detailed model is that the myriad of parameters that might influence excitability, such as specific ion concentrations, are considered to be lumped into the connectivity kernel coefficients that effectively describe the net system gain. Analytic solutions... A goal of the framework is to create a method to assimilate ...}

\dean{The framework assumes that the model parameters are stationary over the estimation period. This is a reasonable assumption if applying the framework to an evoked-potential experiment over a short time period in a controlled environment. However, it is well known that factors that influence excitability are non-stationary over longer time periods. Therefore, the estimation framework must be altered if it is to be applied to data over extended periods of time. A possible way to achieve this is to use a more sophisticated model.. Alternatively, a simpler method would be to allow for time-varying parameter estimates...  }

\dean{The implementation of the framework has been described a being appropriate for an evoked-potential experimental protocol. Over the time duration of an experimental paradigm the connectivity and synaptic kernel parameters are assumed to be stationary. However, if the method is applied to a continuous recording over a longer time duration it is unlikely that the parameters will be stationary. If fact, it is likely that changes in the parameter values would be of most interest.}

\dean{To implement the algorithm in a real-time scenario the we would assume that the parameters are stationary over a time period (window) where the estimation is performed. The parameter of the previous window can be used to initialize parameters of the next window.}

>>>>

\subsection{Modeling Assumptions}
\dean{(comment 1 and 9 from reviewer 2:) It is an implicit assumption that the neural field model (in equation~\ref{DiscreteTimeModel}) provides an apt description of the cortical dynamics recorded from the subject. Although models of this form are capable of describing a variety of cortical dynamics, there is without doubt, a mismatch between the cortex and the model. Nevertheless, there is a sufficient volume of interesting results from theoretical studies that warrants the data assimilation framework. }

\dean{An advantage in using a mean field model over a more detailed mathematical description is that the myriad of parameters that might influence excitability, such as specific ion concentrations, are considered to be lumped into the connectivity kernel coefficients that effectively describe the net system gain. In this way a minimal number of assumptions are made. A major challenge in data assimilation for the mass action of the brain is to make models sufficiently detailed such that the parameters are meaningful, but simple enough to yield clear insights that can be related to theoretical studies (from neuroscience and engineering perspectives) .}

\dean{The framework assumes that the parameters of the model (and brain) are stationary over the estimation period. This is a reasonable assumption if applying the framework to an evoked-potential experiment over a short time period in a controlled environment. However, it is well known that factors that influence excitability are non-stationary over longer time periods (see \citet{Ullah2010} for a nice example). Therefore, the estimation framework must be altered if it is to be applied to data over extended periods of time. A possible way to achieve this is to use a more sophisticated model that accounts for longer term time dependancies (higher temporal order). Alternatively, the parameters of the model could be considered time-varying, where changes in excitability might result in different connectivity kernel parameters. A possible way to achieve this would be to employ window-based estimation, where the estimation is performed on successive windows of data (possibly overlapping windows). The window length will need to be sufficiently short such that the parameters are approximately stationary of the estimation duration, and sufficiently long to achieve a good estimation result. A real-time implementation can also be achieved using the UKF (without the smoother) and a real-time LS estimator (see REF for example). }
